### 2016-11-28 thesis/intro/intro.tex
tmp commit

% TODO: do I really need tumour heterogeneity??
\subsubsection{Tumour heterogeneity (?)}
\label{subsub:tumour_heterogeneity}

\subsubsection{Cancer stem cells (?)}
\label{ssub:cancer_stem_cells}

@@

@@

p53 mutations, BRCA?

\subsubsection{Environmental factors}
\label{ssub:Environmental factors}

@@

\subsection{Microarray technology}
\label{subsec:microarray_technology}

\subsection{Next generation sequencing technology}
\label{sub:next_generation sequencing_technology}

@@

% TODO: have a leading sentence that introduces the gene signatures

\section{Obesity associated genetic signatures}
\label{sec:obesity_associated_genetic_signatures}

\subsection{\citet{Creighton2012} study}
\label{sub:creighton_study}

\subsection{\citet{Fuentes-Mattei2014} study}
\label{sub:fuentes_mattei_study}

\section{Pathway associated genetic signatures}
\label{sec:pathway_associated_genetic_signatures}

@@

\subsection{\citep{Gatza2011} study}
\label{sub:gatza_study}

>>>>

% TODO: do I really need tumour heterogeneity??
\subsubsection{Tumour heterogeneity (?)}
\label{subsub:tumour_heterogeneity}
TBC

\subsubsection{Cancer stem cells (?)}
\label{ssub:cancer_stem_cells}
TBC

@@

There are many causes and risk factors associated with the development and progression of cancer.
Like obesity, cancers are dependent on both the genetic and environmental factors.

@@

It is quite obvious from the previous section (\cref{sub:mechanisms_of_cancer_development}: \nameref{sub:mechanisms_of_cancer_development}) that there are a lot of things that can go wrong in the cell.
Any one of the ten Hallmarks of Cancers has the potential to initiate cancer growth and progression; every hallmark has their own complex mechanistic pathways; every pathway has many proteins associated with its function or role in the cell; and, though their impacts may differ from protein to protein, mutations in any one of the proteins involved in a pathway has the potential to cause cancer.

% TODO: more citations!!!!
Broadly speaking, there are three major classes of genes that have significance in the initiation of cancer: oncogenes, tumour suppressor genes and stability genes \citep{Vogelstein2004a}.
Oncogenes are genes that promote or ``accelerate'' the process of tumorigenesis and/or tumour progression, whereas tumour suppressor genes prevent or ``decelerate'' those effects \citep{Vogelstein2004a}.
Cancers usually mutate the oncogenes in such a way that it it constitutively activated or expressed at a very high level, and mutate the tumour suppressor genes so that these genes are inactivated and/or have reduced expression \citep{Vogelstein2004a}.
Another thing to note for these genes is that oncogenes usually only require a ``one-hit'' mutation that affects a single allele, whereas tumour suppressor genes require mutations in both maternal and paternal alleles to have any effect \citep{Stratton2009,Vogelstein2004a}.

Stability genes, also known as caretakers, are genes that are involved in, for example,  \gls{mmr}, \gls{ner} and \gls{ber}; in other words, genes that are involved in \acrshort{dna} repair mechanisms \citep{Vogelstein2004a}.
These caretaker genes in effect prevents the cell to develop new mutations from \acrshort{dna} replication and mitosis.
However, only the oncogenes or the tumour suppressor genes are able to directly affect the biology of the cell, and therefore mutations in the stability genes affect the mutation rates of the other classes of genes, ultimately leading to tumorigenesis \citep{Vogelstein2004a}.
\\

\noindent
There are two types of mutations that occurs in an individual: germline mutation and somatic mutation.
Germline mutations are the mutations that are passed on from your parents, and somatic mutations are the mutations acquired by the cell over the course of the individual's lifetime.

Unfortunately, there are germline mutations that make the individual more prediposed to cancer \citep{Vogelstein2004a}.
These individuals have a mutation in one or more of the classes of cancer-related genes, which makes them more likely to develop tumour than their normal pairs \citep{Vogelstein2004a}.
As a consequence, these individuals may develop multiple tumours at an early stage of their lives than those who develop tumours from somatic mutations alone, as these individuals already have a mutation that allows all of the cells in their body to have one or more of the Hallmarks of Cancers \citep{Vogelstein2004a}.

In contrast, somatic mutations are acquired during an individual's lifetime and is not passed down to their children.
Somatic mutations can either be a driver or a passenger mutation, where a driver mutation helps tumorigenesis and its subsequent progression (in other words, mutations in oncogene, tumour suppressor, or stability gene), and passenger mutation is where the mutation has no selective advantage for the cancer and does not contribute to tumour progression \citep{Stratton2009}.
Since cancers require multiple driver mutations, somatic mutations are an essential part of tumour development.
Perhaps that is one of the reasons why, in some cancers, germline mutation in the stability genes cause greater susceptibility to cancers than mutations in oncogene or tumour suppressor gene \citep{Vogelstein2004a}.

\subsubsection{Environmental factors}
\label{ssub:environmental_factors}

Though mutation rates can differ between individuals,

``The risk of lung cancer diminishes after smoking cessation, but not during the first 5 years, and the relative risk never turns to that of a nonsmoker'' \citep{Hecht1999}. \\

Tobacco carcinogens form DNA adducts (carcinogen metabolite attached covalently to DNA at A or G), which can be activated or detoxfied in the cell \citep{Hecht1999}. \\

If DNA adducts are not detoxfied, it can lead to DNA miscoding and therefore mutation \citep{Hecht1999}. \\

@@

\subsection{Microarray and Next Generation Sequencing technologies}
\label{sub:microarray_and_next_generation_sequencing_technologies}

@@

\subsection{Obesity associated genetic signatures}
\label{sub:obesity_associated_genetic_signatures}

\subsubsection{\citet{Creighton2012} study}
\label{ssub:creighton_study}

\subsubsection{\citet{Fuentes-Mattei2014} study}
\label{ssub:fuentes_mattei_study}

\subsection{Pathway associated genetic signatures}
\label{sub:pathway_associated_genetic_signatures}

@@

\subsubsection{\citet{Gatza2011} study}
\label{ssub:gatza_study}
